Analysis of cloud enhancement events in a 30-year record of global solar irradiance at Thessaloniki, Greece

### 2024-09-08 article/article.Rmd
Commit 2024-09-08 12:14

In this study, we investigate the characteristics of global horizontal irradiance
enhancement events induced by clouds over Thessaloniki for the period 1994 -- 2023
using data recorded every one minute.  We identified the cloud enhancement (CE)
events by creating an appropriate cloud-free irradiance reference using a radiative
transfer model and aerosol optical depth data from a collocated Cimel sun photometer
and a Brewer spectrophotometer. We found a trend in CE events of
$`r plus_sign(round(yearlyCE[var=="GLB_diff.N",slope],0))`\pm
`r round(2*yearlyCE[var=="GLB_diff.N",slope.sd],0)`\,\text{cases}/\text{year}$,
and a trend in the corresponding irradiation of
$`r plus_sign(round(yearlyCE[var=="GLB_diff.sum",slope],1))`\pm
`r round(2*yearlyCE[var=="GLB_diff.sum",slope.sd],1)`\,\text{kJ}/\text{year}$.
The peak of the CE events was observed during May and June. CE events with duration
longer than 10\nobreakspace{}min are very rare
($<`r ceiling(100*ST_G0[,sum(GLB_ench.N>10)/.N])`\,\%$), with exceptions lasting
over an hour and up to 140 minutes.  Finally, we have detected enhancements above the
total solar irradiance at the top of the atmosphere for the same solar zenith angle
of up to
$400\,\text{W}/\text{m}^{2}$, with the $75\,\%$ of the cases below
$200\,\text{W}/\text{m}^{2}$. Most of these  extreme events occur in spring -- early
summer, with a secondary peak in autumn.

@@

For the detection of CE cases, we established a baseline of irradiance above which we
characterized each data point as an enhancement event and calculated the over
irradiance (OI). The OI is defined as the irradiance difference of the measured
one-minute GHI from the $\text{GHI}_\text{ref}$ corresponding to cloud-free
atmosphere.  First, we used a simple approach for the determination of
$\text{GHI}_\text{ref}$: The Haurwitz's model [@Haurwitz1945], which is a simple
clear sky radiation model and was already adjusted and applied to our data
[@Natsis2023].  We created a threshold by using an appropriate relative and/or an
additional constant offset.  The initial results showed that we can detect a big
portion of the actual CE events.  However, by inspecting the daily plots of
irradiance it became evident that changes in atmospheric conditions introduced
numerous false positive or false negative results. The main reason for these
discrepancies is the variability of the effects of aerosols and water vapor which
were not considered in the simple method.  To produce a more representative reference
we included the effects of these factors using a radiative transfer model (RTM).  The
applied methodology is discussed in section\nobreakspace{}\ref{rtmcs}.

@@

>>>>

In this study, we investigate the characteristics of global horizontal irradiance
enhancement events induced by clouds over Thessaloniki for the period 1994 -- 2023
using data recorded every one minute.  We identified the cloud enhancement (CE)
events by creating an appropriate cloud-free irradiance reference using a radiative
transfer model and aerosol optical depth data from a collocated Cimel sun photometer
and a Brewer spectrophotometer. We found a trend in CE events of
$`r plus_sign(round(yearlyCE[var=="GLB_diff.N",slope],0))`\pm
`r round(2*yearlyCE[var=="GLB_diff.N",slope.sd],0)`\,\text{cases}/\text{year}$,
and a trend in the corresponding irradiation of
$`r plus_sign(round(yearlyCE[var=="GLB_diff.sum",slope],1))`\pm
`r round(2*yearlyCE[var=="GLB_diff.sum",slope.sd],1)`\,\text{kJ}/\text{year}$.
The peak of the CE events was observed during May and June. CE events with duration
longer than 10\nobreakspace{}min are very rare
($<`r ceiling(100*ST_G0[,sum(GLB_ench.N>10)/.N])`\,\%$), with exceptions lasting
over an hour and up to 140 minutes.  Finally, we have detected enhancements above the
total solar irradiance at the top of the atmosphere for the same solar zenith angle
of up to
$209\,\text{W}/\text{m}^{2}$, with the $75\,\%$ of the cases below
$40\,\text{W}/\text{m}^{2}$. Most of these extreme events occur in spring -- early
summer, with a secondary peak in autumn.

@@

For the detection of CE cases, we established a baseline of irradiance above which we
characterized each data point as an enhancement event and calculated the over
irradiance (OI). The OI is defined as the irradiance difference of the measured
one-minute GHI from the $\text{GHI}_\text{ref}$ corresponding to cloud-free
atmosphere.  First, we used a simple approach for the determination of
$\text{GHI}_\text{ref}$: The Haurwitz's model [@Haurwitz1945], which is a simple
clear sky radiation model and was already adjusted and applied to our data
[@Natsis2023].  We created a threshold by using an appropriate relative and/or an
additional constant offset.  The initial results showed that we can detect a big
portion of the actual CE events.  However, by inspecting the daily plots of
irradiance, it became evident that changes in atmospheric conditions introduced
numerous false positive or false negative results. The main reason for these
discrepancies is the variability of the effects of aerosols and water vapor which
were not considered in the simple method.  To produce a more representative reference
we included the effects of these factors using a radiative transfer model (RTM).  The
applied methodology is discussed in section\nobreakspace{}\ref{rtmcs}.

@@

The variable parameters of LUT consists of the month along with the relevant seasonal atmospheric profile,
the climatological values of the alpha, beta parameters of the Ångström
equation for 500nm derived from the 340nm, the WC and the SZA of the simulation.
